Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-07 main.tex
CP
First complete draft of section I

% Notational conventions
\newcommand{\HIGH}{\textsc{h}}
\newcommand{\LOW}{\textsc{l}}
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\PCname(#1)}
\newcommand{\PCname}{\textsc{pc}}
\newcommand{\polname}{P}
\newcommand{\pol}[1]{\polname{#1}}
\newcommand{\trace}[1]{\overline{#1}}
\newcommand{\topo}{T}
\newcommand{\stepsto}{\Rightarrow}
\newcommand{\manystepsto}{\stepsto^\star}

@@

Conventions:
%
\begin{itemize}
\item Words / addresses: $\word,\addr \in \WORDS$
\item Registers: $\reg \in \REGS$
\item Machine states:
$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
\WORDS)$
\item Abbreviations:
\\ $\mach(\addr) = \mach.1(\addr)$
\\ $\mach(\reg) = \mach.2(\reg)$
\\ $\mach(\textsc{pc}) = \mach.3$
\item Step function\footnote{\bcp{I'm taking the step function to be total;
    we could make it partial if we want, but I'm not sure this refinement is
    important.}}: $\mathord{\stepsto} \in \MACHS \rightarrow \MACHS$
\end{itemize}

A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
$\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.

@@

The {\em result} of a machine state $\mach$ (if it exists) is a state
$\mach_n$ such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step
any further.

@@

A {\em topography}\bcp{contour!}
$\topo \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
map from machine components to ``security levels,'' where $\HIGH$ marks
locations containing secret or low-integrity data and $\LOW$ marks those
with public or high-integrity data.

Given a machine state $\mach$ and a topography $\topo$, a {\em variant}
of $\mach$ is any machine state $\mach'$ whose components agree with $\mach$
on all components marked $\LOW$.  \iftext\bcp{Spell it out formally.}\fi

Formally, a {\em policy} $\polname$\iftext\bcp{show its type? then we need
  some more metavariables... :-(}\fi is a function from topographies to
prefix-closed set of traces.  I.e., given a topgraphy, it yields a
restriction of the set of all traces generated by the standard step
function, where some traces may be ``truncated early'' because they violate
the conditions imposed by this topography.

\bcp{Needs to be wrt a policy!!}
Variant machine states $\mach$ and $\mach'$ are said to be {\em
  indistinguishable} with respect to topography $\topo$ and policy
$\pol{\topo}$ if either (1)
one or both of $\mach$ and $\mach'$ fails to terminate, or (2) both
terminate and their results are variants with respect to $\topo$.

A policy $P$

\begin{itemize}
\item Our definition.

\item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
  have to begin with EENI and then get to SSNI as a (stronger) refinement.}

@@

\item Fact: The standard RISC-V semantics does not guarantee NI.  (Give a
little example of a program that reads from a secret location and writes to
a public one, just to illustrate the definition of NI.)
\end{itemize}

>>>>

% Notational conventions
\newcommand{\HIGH}{\textsc{h}}
\newcommand{\LOW}{\textsc{l}}
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\PCname(#1)}
\newcommand{\PCname}{\textsc{pc}}
\newcommand{\polname}{P}
\newcommand{\pol}[1]{\polname(#1)}
% \newcommand{\trace}[1]{\overline{#1}}
\newcommand{\contour}{C}
\newcommand{\CONTOURS}{{\mathcal C}}
\newcommand{\stepsto}{\Longrightarrow}
\newcommand{\stepstounder}[1]{\stackrel{\mbox{\tiny{$#1$}}}{\Longrightarrow}}
\newcommand{\stepstounderfull}{\stepstounder{\textsc{full}}}
\newcommand{\manystepsto}{\stepsto^\star}

@@

Conventions:
%
\begin{itemize}
\item Words / addresses: $\word,\addr \in \WORDS$
\item Registers: $\reg \in \REGS$
\item Machine states:
$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
\WORDS)$
\item Abbreviations:
\\ $\mach(\addr) = \mach.1(\addr)$
\\ $\mach(\reg) = \mach.2(\reg)$
\\ $\mach(\textsc{pc}) = \mach.3$
\end{itemize}

@@

The {\em result} of a machine state $\mach$ with respect to a step function
$\mathord{\stepsto} \in \MACHS \rightharpoonup \MACHS$ is a state $\mach_n$
such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step any
further.

The {\em full} step function, $\mathord{\stepstounderfull} \in \MACHS
\rightharpoonup \MACHS$, embodies the standard RISC-V ISA semantics.

% \bcp{Needed?}A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
% sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
% $\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.

@@

A {\em contour}
$\contour \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
map from machine components to ``security levels,'' where $\HIGH$ marks
locations containing secret or low-integrity data and $\LOW$ marks those
with public or high-integrity data.

Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
components agree with $\mach$ on all components marked $\LOW$ by $\contour$.
\iftext\bcp{Spell it out formally.}\fi

Formally, a {\em policy}
$\polname \in \CONTOURS \rightarrow (\MACHS \rightarrow \MACHS)$ is a
function from contours to step functions (partial functions on $\MACHS$)
that, wherever they are defined, agree with $\stepstounderfull$.  That is,
it represents an enforcement mechanism that sometimes halts programs earlier
than the bare step relation does, but that agrees with the bare step
relation when it allows evaluation to proceed.  We write
$\mach_1 \stepstounder{\pol{\contour}} \mach_2$ for
$\pol{\contour}(\mach_1) = \mach_2$.

If $\mach$ and $\mach'$ are variant machine states with respect to contour
$\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
under policy $\polname$ if either (1) one or both of $\mach$ and $\mach'$
fail to terminate when executed with $\stepstounder{\pol{\contour}}$, or (2)
both terminate and their results are variants with respect to $\contour$.

Policy $\polname$ {\em guarantees noninterference} if, for every $\contour$,
every pair $(\mach,\mach')$ of variant machine states with respect to
$\contour$ are indistinguishable under $\polname$.

\iftext
\begin{itemize}
\item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
  have to begin with EENI and then get to SSNI as a (stronger) refinement.}

@@

\item Fact: The standard RISC-V semantics does not guarantee NI.  (Give a
little example of a program that reads from a secret location and writes to
a public one, just to illustrate the definition of NI.)
\end{itemize}
\fi
